Supplementary Materials

### 2022-12-08 manuscript/supplementary.tex
new supp stuff

>>>>

Our underlying data for all likelihood and pairwise diversity estimates is the
allele count matrix $\mathbf{C}$ with dimensions $l \times 2$, where $l$ is the
chromosome length. This is converted to a pairwise summary matrix with
identical dimensions, $\mathbf{P}$. The first column of $\mathbf{P}$ is the
number of pairwise comparisons between chromosomes that are identical, and the
second column is the number that are different. Both of these columns are
combinatoric summaries of the allele counts matrix. Let $[c_1, c_2]$ be a row
of $\mathbf{C}$ for basepair $i$, and $n = c_1 + c_2$. Then, the (1) total
number of pairwise combinations of chromosomes $n_T$, (2) the number of
pairwise with identical alleles $n_S$, and (3) the number of pairwise
combinations with differing alleles $n_D$ are respectively,

\begin{align*}
  n_T &= \frac{n(n-1)}{2} \\
  n_S &= {c_1 \choose 2} + {c_2 \choose 2} \\
  n_D &= n_T - n_S
\end{align*}

Thus $\mathbf{P}_i = [n_S, n_D]$, and the pairwise diversity is calculated as

\begin{align}
  \bar{\pi}(i) = \frac{n_D}{n_T}
\end{align}

which is identical to the more common expression of this estimator,

\begin{align}
  \bar{\pi}(i) = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
\end{align}

where $k_{i,j}$ is 1 if the alleles at this site differ, and 0 otherwise. Note
that for sites with missing data, $c_1 = c_2 = 0$ and thus $n_T = 0$ and the
division is invalid. Such cases were marked as missing with floating point
values \texttt{NaN}s. Binned summaries of data are weighed by the number of
complete cases, e.g. rows with $c_1 + c_2 > 0$.
